Evolution of pre-copulatory and post-copulatory inbreeding behaviour and associated

### 2016-12-10 ms_pre_post.Rmd
Methods edits up through fertilisation

Phenotypic values for tendency for polyandry ($P_{p}$), pre-copulatory inbreeding strategy ($M_{p}$), and post-copulatory inbreeding strategy ($F_{p}$) each have costs that are assumed to independently increase the probability that a focal female will die before mating. The probabilities of pre-mating mortality due to the costs of polyandry ($c_{P}$), pre-copulatory inbreeding strategy ($c_{M}$), and post-copulatory inbreeding strategy ($c_{F}$) are $P_{p} \times c_{P}$, $|M_{p}| \times c_{M}$, and $|F_{p}| \times c_{F}$, repectively. Here $|M_{p}|$ and $|F_{p}|$ are the absolute values of $M_{p}$ and $F_{p}$, respectively. Absolute values are used for applying costs to inbreeding strategies as appropriate because both negative and positive $M_{p}$ and $F_{p}$ values affect the degree of inbreeding. In contrast, only positive values of $P_{p}$ cause females to be polyandrous (see below). Overall, because generations are non-overlapping, a female's probability of total reproductive failure increases as a linear function with each trait value.

@@

After costs are paid, each remaining female chooses $N_{males}$ males to mate with, where $N_{males}$ is calculated by sampling from a Poisson distribution such that $N_{males} = Poisson(P_{p}) + 1$. This ensures that all surviving females choose at least one mate, and causes some stochastic variation around the expected mean $N_{males}$ of $P_{p}+1$. A female's value of $P_{p}$ depends on the sum of her 20 $P_{a}$ allele values, which might be negative or positive. However, because females cannot mate with a negative number of males, $P_{p}$ is constrained to be zero or positive such that $P_{p} = 0$ when $P_{g} < 0$, but $P_{p} = P_{g}$ when $P_{g} \geq 0$. Polyandry is therefore modelled as a threshold trait [@Lynch1998; @Roff1996; @Roff1998; @Duthie] that is influenced by continuous genetic variation but only expressed at the threshold value of $P_{g} > 0$. Any negative $P_{g}$ value therefore generates phenotypic monandry, while positive $P_{g}$ values can lead to different degrees of realised polyandry (i.e., different numbers of mates, $N_{males}$).

Mating
----------------------------------------------

@@

Most often, $N_{males}$ will be smaller than the total number of available males [@Duthie]. Here, each female chooses her $N_{males}$ mates based on her pre-copulatory inbreeding strategy phenotype ($M_{p}$), which equals her pre-copulatory inbreeding strategy genotype ($M_{g}$) defined as the sum of the values of her 20 $M_{a}$ alleles. Negative or positive $M_{p}$ values cause a female to avoid or prefer mating with kin, respectively. Values of $M_{p}=0$ cause females to mate randomly with respect to kinship.

The probability that a focal female $i$ mates with a male $j$ with which she shares some kinship $k_{ij}$ is calculated by first assigning each male a perceived mate quality $Q^{m}_{ij}$. If the female has a strategy of pre-copulatory inbreeding avoidance ($M_{p}<0$), then $Q^{m}_{ij} = (-M_{p} \times k_{ij} + 1)^{-1}$, meaning that $Q^{m}_{ij}$ decreases linearly with increasingly positive values of $k_{ij}$ and increasingly negative values of $M_{p}$. If the female has a strategy of pre-copulatory inbreeding preference ($M_{p}>0$), then $Q^{m}_{ij} = M_{p} \times k_{ij} + 1$, meaning that $Q^{m}_{ij}$ increases with increasingly positive $k_{ij}$ and $M_{p}$. If $M_{p}=0$, then all males are assigned a quality of 1

After a focal female assigns all males a $Q^{m}_{ij}$ value, each male's value is divided by the sum of all $Q^{m}_{ij}$ values across all males, thereby assigning each male a relative perceived quality $q^{m}_{ij}$, which is constrained to values between zero and one. The value of $q^{m}_{ij}$ then defines the probability that a focal female mates with the male; mating is therefore stochastic, and females do not always mate with the male of the highest $q^{m}_{ij}$. For polyandrous females that choose multiple mates (i.e., $N_{mates}>1$), mates are chosen iteratively such that $Q^{m}_{ij}$ and $q^{m}_{ij}$ are re-calculated for each additional mate choice, and with $Q^{m}_{ij}$ and therefore $q^{m}_{ij}$ values of already chosen males set to zero to ensure mate sampling without replacement. In the unlikely event that a female's $N_{males}$ exceeds the total number of available males, then she simply mates with all males.

Fertilisation
----------------------------------------------

>>>>

Phenotypic values for tendency for polyandry ($P_{p}$), pre-copulatory inbreeding strategy ($M_{p}$), and post-copulatory inbreeding strategy ($F_{p}$) each have costs that are assumed to independently increase the probability that a focal female will die before mating. The probabilities of pre-mating mortality due to the costs of polyandry ($c_{P}$), pre-copulatory inbreeding strategy ($c_{M}$), and post-copulatory inbreeding strategy ($c_{F}$) are $P_{p} \times c_{P}$, $|M_{p}| \times c_{M}$, and $|F_{p}| \times c_{F}$, repectively. Here $|M_{p}|$ and $|F_{p}|$ are the absolute values of $M_{p}$ and $F_{p}$, respectively. Absolute values are used for applying costs to inbreeding strategies because both negative and positive $M_{p}$ and $F_{p}$ values affect the degree of inbreeding. In contrast, only positive values of $P_{p}$ cause females to be polyandrous (see below). Overall, because generations are non-overlapping, a female's probability of total reproductive failure increases linearly with each trait value.

@@

After costs are paid, each remaining female chooses $N_{males}$ males to mate with, where $N_{males}$ is calculated by sampling from a Poisson distribution such that $N_{males} = Poisson(P_{p}) + 1$. This ensures that all surviving females choose at least one mate, and causes some stochastic variation around the expected mean $N_{males}$ of $P_{p}+1$. A female's value of $P_{p}$ depends on the sum of her 20 $P_{a}$ allele values (i.e., $P_{g}$), which might be negative or positive. However, because females cannot mate with a negative number of males, $P_{p}$ is constrained to be zero or positive such that $P_{p} = 0$ when $P_{g} < 0$, but $P_{p} = P_{g}$ when $P_{g} \geq 0$. Polyandry is therefore modelled as a threshold trait [@Lynch1998; @Roff1996; @Roff1998; @Duthie] that is influenced by continuous genetic variation but only expressed when $P_{g} > 0$. Any negative $P_{g}$ value therefore generates phenotypic monandry, while positive $P_{g}$ values can lead to different degrees of realised polyandry (i.e., different numbers of mates, $N_{males}$).

Mating and pre-copulatory inbreeding avoidance
----------------------------------------------

@@

Most often, $N_{males}$ will be smaller than the total number of available males [@Duthie]. Each female then chooses her $N_{males}$ mates based on her pre-copulatory inbreeding avoidance phenotype ($M_{p}$), which equals her genotype ($M_{g}$). Negative or positive $M_{p}$ values cause a female to avoid or prefer mating with kin, respectively, whereas $M_{p}=0$ cause females to mate randomly with respect to kinship.

To calculate the probability that a focal female $i$ mates with a male $j$ to whom she is related by some kinship $k_{ij}$, each male is first assigned a perceived mate quality $Q^{m}_{ij}$. If the female has a strategy of pre-copulatory inbreeding avoidance ($M_{p}<0$), then $Q^{m}_{ij} = (-M_{p} \times k_{ij} + 1)^{-1}$, meaning that $Q^{m}_{ij}$ decreases linearly with increasingly positive values of $k_{ij}$ and increasingly negative values of $M_{p}$. If the female has a strategy of pre-copulatory inbreeding preference ($M_{p}>0$), then $Q^{m}_{ij} = M_{p} \times k_{ij} + 1$, meaning that $Q^{m}_{ij}$ increases with increasingly positive $k_{ij}$ and $M_{p}$. If $M_{p}=0$, then all males are assigned $Q^{m}_{ij}=1$.

Each male's value with respect to each focal female is then divided by the sum of all $Q^{m}_{ij}$ values across all males with respect to that female, thereby assigning each male a relative perceived quality $q^{m}_{ij}$, which is constrained to values between zero and one. The value of $q^{m}_{ij}$ then defines the probability that a focal female mates with the male. Mating is realised by placing each $q^{m}_{ij}$ value into a vector whose elements sum to unity, and a random uniform number is then generated to sample a single male from this vector. Mating is therefore stochastic, and females do not always mate with the male of the highest $q^{m}_{ij}$. For polyandrous females that choose multiple mates (i.e., $N_{mates}>1$), mates are chosen iteratively such that $Q^{m}_{ij}$ and $q^{m}_{ij}$ are re-calculated for each additional mate choice, and with $Q^{m}_{ij}$ and therefore $q^{m}_{ij}$ values of already chosen males set to zero to ensure mate sampling without replacement. In the unlikely event that a female's $N_{males}$ exceeds the total number of available males, then she simply mates with all males.

Fertilisation and post-copulatory inbreeding avoidance
----------------------------------------------
